Poseidon Adventure

### 2019-11-22 Sardines_atlantisom2SS_demo.Rmd

Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_22".

We will go from atlantis model output to SS input in this page for a single species, "sardine", that doesn't require any age class splitting.

As of 31 May 2019, the `run_truth` function has been modified to do the catch numbers fix for legacy models.

@@

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

species_ss <- c("Pacific_sardine", "Mesopel_M_Fish")

#function to make a config file? need one for each atlantis run

if(initCCA) source(here("config/CC3Config.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAConfig.R"))

@@

Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions.

## Truth from OM

@@

Plot true biomass and numbers for both "hake" and "sardines."

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% c("Mesopel_M_Fish", "Pacific_sardine"))

@@

Get true population parameters (M in progress, S-R, growth)

```{r true-poppars}

@@

Do true catch numbers times weight at age match catch in biomass? It was about 16 times too high relative to catch.txt output, and shifted one year back (which is due to the floor function, and an easy fix).

June 6: just corrected the catch nums correction in `run_truth` and I hope it matches now.

```{r catchNtoB}

totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard_hake.rds")))

totcatchann <- totcatchN %>%
  mutate(yr = floor(time/fstepperyr)+1) %>%
  group_by(species, yr) %>%
  summarise(anncatchage = sum(atoutput))

plottotC <- ggplot(totcatchann, aes(x=yr, y=anncatchage)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste("Catch in N ", scenario.name,
                        totcatchann$species))

plottotC

catch_lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard_hake.rds")))

CwtAtAge <- catch_lengthwt_census_ss$muweight %>%
  select(species, agecl, time, wtAtAge = atoutput) %>%
  mutate(wtAtAge = wtAtAge/1000)

catchbioN_ss <- merge(catch_numssshigh,CwtAtAge,by=c("species","agecl", "time"),all.x=T) %>%
  mutate(biomass = atoutput*wtAtAge/1000) %>%
  mutate(yr = floor(time/fstepperyr)+1) %>%
  group_by(species, yr) %>%
  summarise(anncatchB = sum(biomass))

plotcatch <- ggplot() +
  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio from txt"),
            alpha = 10/10) +
  geom_point(data=catchbioN_ss, aes(x=yr,y=anncatchB, color="true catch bio from N"),
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

Sadly, still a problem with the catch correction if it should match catch.txt when converted to biomass. We are now off by about a factor of 2 (catch nums * wtAtAge is half true catch bio from the txt file), but this differs slightly by species:

```{r catchoff}

 sardcatchB <- truecatchbio_ss #%>%
# now done above
#   mutate(yr = as.integer(time/365))

sardcatchBcomp <- merge(sardcatchB, catchbioN_ss)
sardcatchBcomp <- mutate(sardcatchBcomp, mult=atoutput/anncatchB)

#mean(sardcatchBcomp$mult)

#summary(sardcatchBcomp$mult)
Filter(Negate(is.null), tapply(sardcatchBcomp$mult, sardcatchBcomp$species, summary))

## “Data” from OM for EMs

@@

Now we are going single species. First we generate the inputs for sardine.

@@

Get input total catch for SS (here using numbers). SS takes a fishery cv which can be applied here to get a catch time series, or we can input true catch.

PLEASE DON'T USE THIS FOR CCA. We have determined above that catch in numbers from legacy atlantis code, even corrected, is still off by a factor of 2. We leave the info here because this code can be used for atlantis codebases after December 2015.

```{r toSS-fisheryts}

# catch at age by area and timestep
catch_numbers <-  create_fishery_subset(dat = truth$catch,
                                         time = fish_times,
                                         species = species,
                                         boxes = boxall)

if (fstepperyr>1){
  #need to take all output time steps and sum catch over each year
  catch_numbers_area <- catch_numbers %>%
    #assign a temporary year column, aggregate nums over this year
    mutate(yr = floor(time/fstepperyr)+1) %>%
    group_by(species, agecl, polygon, yr) %>%
    summarise(anncatch_area=sum(atoutput))

  # catch in numbers summed over polygon 
  # and agecl after time has been aggregated into 1 yr
  catch_total <- catch_numbers_area %>%
    group_by(yr) %>%
    summarise(catch=sum(anncatch_area))
}

# catch in numbers summed over polygon and agecl if output is annual
if (fstepperyr==1){
  catch_total <- catch_numbers %>%
    group_by(time) %>%
    summarise(catch=sum(atoutput))
}

# if we want to apply the CV to this we need to write sample_total_catch

#cv <- data.frame(species="Pacific_sardine", cv=CVs$fishery)

#catchObsNum <- sample_total_catch(dat=catch_numbers, cv=cv)

Check this against true total catch numbers above. Should be the same (and it is, but both are incorrect in the absolute scale due to the CATCH.nc output problem in older codebases).

```{r compare-totcatch}

compareC <- ggplot() +
  geom_line(data=totcatchann, aes(x=yr, y=anncatchage, color="census catch N"), 
            alpha = 10/10) +
  geom_point(data=catch_total, aes(x=yr,y=catch, color="sample catch N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareC +
  facet_wrap(~species, scales="free")

@@

#catch at age each timestep summed over polygons
catch_numsss_samp <- sample_fish(catch_numbers, effN)

@@

<!--Can we match catch in weight from catch.txt? No, numbers sampled are 1e-6 compared with actual catch numbers.

```{r compareC-wt, eval=FALSE}

CswtAtAge <- catch_lengthwt_samp$muweight %>%
  select(species, agecl, time, wtAtAge = atoutput) %>%
  mutate(wtAtAge = wtAtAge/1000)

catchsampbio_ss <- merge(catch_numsss_samp,CswtAtAge,by=c("species","agecl", "time"),all.x=T) %>%
  mutate(biomass = atoutput*wtAtAge/1000) %>%
  mutate(yr = floor(time/fstepperyr)+1) %>%
  group_by(species, yr) %>%
  summarise(anncatchBsamp = sum(biomass))

plotcatch <- ggplot() +
  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput/1000000, color="true catch bio"),
            alpha = 10/10) +
  geom_point(data=catchsampbio_ss, aes(x=yr,y=anncatchBsamp, color="sample catch B"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

```
-->

@@

Write time series biomass (tons) and catch (tons) data for SS using `SS_write_ts`:

@@

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1 # this is not reading in, Lbin_method is name, but synthesis can't run if this changes from the original 3 to a 1, so I'm leaving it
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

>>>>

This demonstrates the Atlantis to `atlantisom` to SS3 workflow with the California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_22".

We will go from Atlantis model output to SS input and run in this page for a single species, "sardine", that doesn't require any age class splitting.

@@

species_ss <- c("Pacific_sardine")

source(here("config/CC3Config.R"))

@@

Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions. Here the survey functions are still separate.

## Survey Census

@@

Plot true biomass and numbers for "sardines."

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% c("Pacific_sardine"))

@@

@@

## Derive “Data” from Atlantis to give to SS3

@@

@@

@@

#catch at age each timestep summed over polygons
# catch at age by area and timestep
catch_numbers <-  create_fishery_subset(dat = truth$catch,
                                         time = fish_times,
                                         species = species,
                                         boxes = boxall)

catch_numsss_samp <- sample_fish(catch_numbers, effN)

@@

@@

Write time series biomass (tons) and catch (tons--taken directly from truth without error) data for SS using `SS_write_ts`:

@@

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

### 2019-11-22 docs/index.Rmd
add initial comparisons

@@

>>>>

### Demo for NEMoW meeting, December 2019

<a href="https://sgaichas.github.io/poseidon-dev/Sardines_atlantisom2SS_demo.html" target="_blank">Sardines in the CalCurrent run with F, rec dev, and climate scenario, all true and SS data, runs SS3, November 21, 2019</a>

@@

So we can start some comparisons of true and model estimated attributes.

<a href="https://sgaichas.github.io/poseidon-dev/SkillAssessInit.html" target="_blank">Compare Atlantis truth with SS3 estimated B, R, and F, October 1, 2019</a>
